Dish: Distribution-oblivious Shell Scripting

### 2019-11-22 pldi/draft.tex
Pass over extensibility

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_schematic.pdf}}
\caption{
  \textbf{\sys transformation overview.}
  Given a shell pipeline, \sys leverages command distributability to identify
  high-distributability stages, rewrite them, and orchestrate their execution.
}
\label{fig:overview}
\end{figure}

@@

Given a set of commands composed together by the shell primitives, \sys's next challenge is to analyze it and identify candidate subexpressions for distribution~\sx{ir}.
\sys addresses this challenge with a rewriting \todo{pass} that performs dataflow analysis identifying \todo{maximal} distributable sub-expressions.
It converts the AST to a dataflow graph, and iteratively performs graph transformations that expose data parallelism as task parallelism while preserving correctness.
These transformations are complicated by the shell's pervasive use of dynamic features---\eg environment variables, file input sizes, \etc

Finally, given a distributed pipeline as a dataflow graph, \sys needs to schedule and execute it.
To achieve this, \sys's runtime component needs to solve several challenges~\sx{impl}.
First, \sys needs to wrap single-node commands with corrective runtime interposition based on their classes;
  this is achieved by thin wrappers for all the commands involved in the program, generated just-in-time using the distributability specification outlined earlier.
The reasoning behind wrappers being thin is to not interfere with legacy functionality---commands have many flags accumulated over long periods of time.

@@

\heading{Discussion}
As described earlier~\sx{bg}, shell commands change their operation based on command-line options.
These options can actually change the class in which a command belongs.
Consider two examples:
  (i) \ttt{cat} is in \sta, but with \ttt{-n} it jumps into \pur because it has to keep track of a counter;
  (ii) \ttt{comm} is in \pur, but with \ttt{/input1} it jumps into \sta---as it is now partially evaluated, it becomes stateless with respect to the second argument.
% TODO: closure?

@@

One of the major characteristics of the shell is its extensibility, as
also mentioned in \Cref{bg}. Existing commands are modified and
extended over time and users are free to install custom commands from
a wide variety of sources.
%% \kk{Text point: This restricts the space of solutions for
%%   distribution of shell scripts}
This means that any one-time solution for categorizing shell commands
is not satisfactory, since it would very quickly become obsolete,
being able to support a very limited amount of shell scripts.

%% \kk{Text point: The characteristics of a good solution}

A solution that applies to shell programms should thus be extensible,
in the sense that adding or changing a command, doesn't invalidate the
existing information on command categories. A developer of a new
command, should be able to categorize it accordingly, and succinctly
express the information that the system needs in order to distribute
it.

%% \kk{Solution: A lightweight language that can be used by a command
%%   developer to categorize it for distribution.}

\TODO{Make sure that before this point, we mention that our notion of
  command has one input and one output. Concretely: the abstract
  notion of a program that we have is programs with one input and one
  output stream. The command developer must be able to express the
  input argument of a commmand (or if it is the stdin).  }

We address this issue by designing a lightweight language whose
purpose is to categorize a command and describe the sequence of
sources from which it reads its input.

%% \subsection{Categorization Language}

%% \kk{Text point: Introduce problem: Most shell commands, belong to
%%   different classes depending on their flags, arguments}

\TODO{Mention the point of these two paragraphs in the intro?}

Consider the command \ttt{comm}. It takes as input two files and
implements some form of join. When invoked without any option, it
produces three-column output. The first column contains lines that
unique to the first file, the second column contains lines that are
unique to the second file, and the last column contains lines that
exist in both files. In the general case \ttt{comm} is pure, since it
needs to wait until it reads both files completely before it can
output its results.

However, \ttt{comm} can also be invoked using some options that
supress the output of any combination of columns. More precisely, the
option \ttt{-1} suppresses the first column of the output. Similarly,
\ttt{-2} and \ttt{-3} supress the second and third column. Using both
\ttt{-2} (or \ttt{-1}) and \ttt{-3}, the command becomes stateless
regarding its first (resp. the second) file argument.

@@

    <command-list> ::= <command>
    \alt <command> <command-list>
  \end{grammar}
  \caption{\todo{Categorization language}}
  \label{fig:dsl}
\end{figure}

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_schematic.pdf}}
\caption{
  \textbf{\sys transformation overview.}
  Given a shell pipeline, \sys leverages command distributability to identify
  high-distributability stages, rewrite them, and orchestrate their execution.
}
\vspace{-15pt}
\label{fig:overview}
\end{figure}

@@

Given a set of commands composed together by the shell primitives, \sys's next challenge is to analyze it and identify candidate subexpressions for distribution~\sx{ir}.
\sys addresses this challenge with a series of rewriting passes that perform dataflow analysis for identifying \todo{maximal} distributable sub-expressions.
It converts the AST to a dataflow graph, and iteratively performs graph transformations that expose data parallelism as task parallelism while preserving correctness.
These transformations are complicated by the shell's pervasive use of dynamic features---\eg environment variables, file input sizes, \etc

Finally, \sys needs to schedule and execute the distributed dataflow graph.
To achieve this, \sys's runtime component needs to solve several challenges~\sx{impl}.
First, \sys needs to wrap single-node commands with corrective runtime interposition based on their classes;
  this is achieved by thin wrappers for all the commands involved in the program, generated just-in-time using the distributability specification outlined earlier.
The reasoning behind wrappers being thin is to not interfere with legacy functionality---commands have many flags accumulated over long periods of time.

@@

\heading{Discussion}
There are several technical details about \sys's design decisions worth noting.

In terms of inputs, \sys supports commands with multiple input (resp. output) streams, by prepending (resp. appending) shell built-ins that merge input (resp. split output) streams~\sx{impl}.
Thus, a command can be modelled as having one input and one output stream, a representation that is used throughout \sys including the extensibility DSL~\sx{ext}.

Wanting to present the full picture for completeness purposes, the focus for the rest paper will be commands in \sta and \pur.
Other commands are treated as  non-distributable for now~\sx{conclusion};
  by building infrastructure to handle other classes, \sys would be able to (i) extract more parallelism and (ii) handle larger programs.
Even with only these two classes, however, \sys supports highly useful pipelines that are more expressive than ones in popular frameworks such as Hadoop~\cite{mapreduce:08} and Spark~\cite{spark:12}~\sx{eval}.

@@

% FIXME: \nv{TODO: remember to bubble up(me)}
An important characteristic of the \unix shell is its language-agnostic extensibility~\sx{bg}.
Users composing pipelines are free to install custom commands available from various sources.
Being general programs, commands are developed in a variety of languages and are extended over long periods of time.
%% \kk{Text point: This restricts the space of solutions for
%%   distribution of shell scripts}
As a result, any static characterization of shell commands is unsatisfactory as it would quickly become obsolete, supporting only a small subset of commands.
This poses a challenge to the developers of individual commands---how can they augment their commands to capture information critical to \sys without too much effort?
%% \kk{Text point: The characteristics of a good solution}

To address additions and changes to the set of commands available to the shell, \sys provides command developers lightweight, domain-specific language (DSL).
The DSL allows developers to express semantic information about a command's distributanility that is critical to \sys.
This DSL can be used by the developers of new commands succinctly express this information, such the class of a command and the sequence of sources from which it reads its input.
It is also used by developers that work on existing commands to extend or maintain them, to express additions or changes to the command's semantics.

\heading{Command-line Options}
As described earlier~\sx{bg}, command-line options and environment variables can affect the operation of a command.
Options are a particularly popular way to control a command, 
  directly affecting a command's distributability classification.
That is, commands are assigned to default classes, and commands can change the class in which a command belongs.

As simple examples, consider \ttt{cat} and \ttt{sort}.
By default \ttt{cat} is in \sta, but with \ttt{-n} it jumps into \pur because it has to keep track of a counter.
Conversely, \ttt{sort} defaults to \pur, but \ttt{with} \ttt{-h} it jumps into \sta.
% TODO: closure?

\heading{Command Inputs and Outputs}
Consider the command \ttt{comm}. It takes as input two files and
implements some form of join. When invoked without any option, it
produces three-column output. The first column contains lines that
unique to the first file, the second column contains lines that are
unique to the second file, and the last column contains lines that
exist in both files. In the general case \ttt{comm} is pure, since it
needs to wait until it reads both files completely before it can
output its results.

However, \ttt{comm} can also be invoked using some options that
suppress the output of any combination of columns. More precisely, the
option \ttt{-1} suppresses the first column of the output. Similarly,
\ttt{-2} and \ttt{-3} supress the second and third column. Using both
\ttt{-2} (or \ttt{-1}) and \ttt{-3}, the command becomes stateless
regarding its first (resp. the second) file argument.

% \ttt{comm} is in \pur, but with \ttt{/input1} it jumps into \sta---as it is now partially evaluated, it becomes stateless with respect to the second argument.

@@

    <command-list> ::= <command>
    \alt <command> <command-list>
  \end{grammar}
  \caption{
\textbf{Distributability description language.}
\todo{The DSL attempts to capture important elements regarding the distributability aspects of various commands.}
\vspace{-15pt}
}

  \label{fig:dsl}
\end{figure}
